---
title: "Nuclear weapons devolopment over time"
author: "Olivia Kusk, Natasha Due, Simone Bærnthsen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

### Explore Nuclear Weapons' stockpile over time
We want to use ggplot to investigate how the development of Nuclear Weapons' stockpile in Russia and the US over the period from 1945-2024.
```{r libraries, echo = TRUE}
#running of needed library
library(tidyverse)
library(ggplot2)
library(here)

#getting the data
NW_stockpiles <- read_csv("data/NW_stockpile.csv")
```

### Check data and renaming columns
We are going to check if the data is correctly loaded and if anythings need to be changed to make it easier to work with, and renaming column 1 from "Entity" to "Country" and column 4 from "Number of nuclear warheads" to "Number_of_NW".
```{r datacheck, eval = TRUE}
#Renaming columns
names(NW_stockpiles)[1]<-"Country"
names(NW_stockpiles)[4]<-"Number_of_NW"

#Check the data
str(NW_stockpiles)
unique(NW_stockpiles$Country)
first(NW_stockpiles$Year)
glimpse(NW_stockpiles)
```

### Plotting of the nuclear warhead stockpiles
We are going to create an graph that compare all of the countries with nuclear warheads and the whole world.
```{r plot-nuclear-wearheads, eval = TRUE}
Nw_stockpiles <- NW_stockpiles %>% 
  ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_linedraw()
Nw_stockpiles

ggsave(here("figures","nw_all.png"), plot = Nw_stockpiles, height = 5)
```

### Plotting of the nuclear warhead stockpiles
To better understand the individual among of nuclear warheads pr country and the world, we will also splitting them into individual graphs.
```{r plot-nuclear-warheads-individual, eval = TRUE}
Nw_individual <- NW_stockpiles %>% 
  ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_bw()+
  facet_wrap(~Country)

Nw_individual

ggsave(here("figures","nw_all_individual.png"), plot = Nw_individual, height = 5)
```

### Creating a dataset without Russia and the US
Because Russia and the US (and the whole world) have a lot more nuclear warheads than the rest of the countries, will we create a new dataset without Russia, the US and the World.
```{r creating-new-dataset, eval = TRUE}
Without_Rus_US <- NW_stockpiles %>% 
  filter(Country!="Russia") %>% 
  filter(Country!="United States") %>% 
  filter(Country!="World")
```

### Plotting of the nuclear warhead stockpiles
We are going to create an graph that compare the countries with nuclear warheads after removing Russia, the US and the World.
```{r plot-nuclear-wearheads-without, eval = TRUE}
Graph_W_Rus_US <- Without_Rus_US %>% 
   ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_linedraw()

Graph_W_Rus_US
ggsave(here("figures","nw_without_rus_us.png"), plot = Graph_W_Rus_US, height = 5)
```

### Plotting of the nuclear warhead stockpiles individual
Once again we are splitting the individual graphs to better see the individual countries data.
```{r plot-nuclear-warheads-individual2, eval = TRUE}
Graph_W_Rus_US_indi <- Without_Rus_US %>% 
  ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_bw()+
  facet_wrap(~Country)

Graph_W_Rus_US_indi
ggsave(here("figures","nw_w_rus_us_indi.png"), plot = Graph_W_Rus_US_indi, height = 5)
```

### Creating a dataset only with Russia and the US
We will also create a dataset only with Russia and the US.
```{r creat-nuclear-wearheads-Rus-US, eval = TRUE}
Only_Rus_US <- NW_stockpiles %>% 
  filter(Country =="Russia"|Country=="United States")
```

### Plotting of the nuclear warhead stockpiles
We are going to create an graph that compare the countries with nuclear warheads after removing Russia, the US and the World.
```{r plot-nuclear-wearheads-Rus-Us, eval = TRUE}
Graph_Rus_Us <- Only_Rus_US %>% 
  ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_linedraw()

Graph_Rus_Us
ggsave(here("figures","nw_rus_us.png"), plot = Graph_Rus_Us, height = 5)
```

### ### Plotting of the nuclear warhead stockpiles individual
Here will we split Russia and the US into individual graphs to better understand the countries data.
```{r plot-nuclear-warheads-individual3, eval = TRUE}
Graph_Rus_Us_indi <- Only_Rus_US %>% 
  ggplot(aes(x=Year,y=Number_of_NW, colour = Country))+
  geom_line()+
  theme_bw()+
  facet_wrap(~Country)

Graph_Rus_Us_indi
ggsave(here("figures","nw_rus_us_indi.png"), plot = Graph_Rus_Us_indi, height = 3, width = 6)
```

